Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-08-16 GWLDCII.tex
more comments and corrections, now to the end

Our next goal is to show that the dimension of the positroid cell defined by a Wilson loop diagram $(\cP, [n])$ has dimension $3|\cP|$ (Theorem \ref{thm dim}).  The positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite{non-orientable}, Agarwala and Marcott show that the dimension of the Deodhar component is 
\[|\cP| + \text{ the dimension of associated positroid cell}.\]
In other words, the Deodhar component associated to a Wilson loop diagram is $4|\cP|$, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron.

@@

\begin{proof}
  By Algorithm~\ref{alg:put GN on WLD} the Grassmann necklace of $W$ is obtained from the Grassmann necklace of $V$ by duplicating the $i$th element of the Grassmann necklace of $V$ (shifting indices by $1$ for all indices greater than $i$), and incrementing all indices greater than $i$ by $1$ in each Grassmann necklace element.  Formally, if $I_1^{V}, \ldots, I_{n-1}^{V}$ and $I_1^{W}, \ldots, I_n^{W}$ are the Grassmann necklaces of $V$ and $W$ respectively then
  \[
  I_j^{W} =
  \begin{cases}
    \{\ell \in I_j^{V} : \ell < i\} \cup \{\ell+1 \in I_j^{V} : \ell \geq i\} & \text{if $j\leq i$} \\
    \{\ell \in I_{j-1}^{V} : \ell < i\} \cup \{\ell+1 \in I_{j-1}^{V} : \ell \geq i\} & \text{if $j > i$.}
  \end{cases}
  \]

By Lemma~\ref{no coloops} we know that $i \not\in I_1^W$, and so $i$ must label a horizontal edge on the boundary of the Le diagram of $W$, i.e. it must be a column label. The shapes of the Le diagram of $V$ and $W$ are the same except for the insertion of this column since $I_1^*$ is the same for $V$ and $W$ except for the incrementation of the indices $\geq i$ in the transition from the necklace for $V$ to the necklace for $W$. 
\end{proof}

@@

We first consider $I_1^{W}$, i.e. the set obtained by applying the Grassmann necklace algorithm to $W$ with starting vertex 1. Note that $n-3 \not\in I_1^{V}$: for Case 2 this is clear from the diagram (since Algorithm~\ref{alg:put GN on WLD} would assign $q$ no later than vertex $n-4$), while for Case 1 it follows from Lemma \ref{lem no fourth vertex}. Therefore when we start at vertex 1 and apply the Grassmann necklace algorithm to $W$, $p$ is the only remaining unassigned propagator when we reach vertex $n-3$, and so we have $ I_1^{W}(p) = n-3$ and $I_1^W = I_1^V\cup \{n-3\}$.

@@

Finally, if $n-2,n-1 \in I_{n-2}^{V}$ then $I_{n-2}^W(q) = n-2$ and $I_{n-2}^W(p) = n-1$ as above, but this is different to what occured in $I_{n-2}^{V}$ so we cannot use the same argument as above. When the algorithm $I_{n-2}^W$ reaches vertex $n$, the only propagators $p$ and $q$ have contributed; thus we proceed as in the construction of $I_n^{V}$ but without propagator $q$. By Lemma~\ref{vertex cyclic int lem} we know that $I_n^V(q) = n-5$, and from the diagram we see that the only way this could occur is if all other propagators in $V$ had already been assigned when we reached vertex $n-5$; thus $q$ was the final propagator to be assigned in $I_n^{V}$. Therefore $I_{n-2}^{W} = (I_{n}^{V} \setminus \{n-5\}) \cup \{n-1, n-2\}$ in this case. This completes all cases for $I_{n-2}^{W}$.

@@

As illustrated in Figure~\ref{fig Le}, the pieces of the Le diagrams of $V$ and $W$ will be called $\mathcal{A}$ and $\mathcal{B}$ in what follows. \hlfix{Explicitly, $\mathcal{B}$ is the set of columns with labels $v < n-3$ and $\mathcal{A}$ is the set of columns labeled by $n$, $n-1$ and $n-2$ in $V$. In $W$, $\mathcal{A}$ consists of these columns without the last row.}{is this accurate?} Over the course of the next few lemmas we will prove that the plusses in the $\mathcal{B}$ parts of each diagram are identical, and that the relationship between the plusses in the $\mathcal{A}$ regions can be described explicitly.

@@

Finally, we also note that when we speak of a plus in the Le diagram of $V$ being the same as in $W$ or vice versa, we mean that the position of the plus in $\mathcal{A}$ or $\mathcal{B}$ is the same; because of the column insertion the absolute indices may differ.

@@

In this section, we investigate the poles of the integrand for $I(W)$, that is $R(W)$ as defined in Definition~\ref{def R(W)}. It is conjectured that for a fixed number of propagators and a given set of external twistors $\cZ$, [cite] \bas \sum_{W | |\cP| = k} I(W)(\cZ) <\infty \;.\eas In otherwords, all poles of the intergrals must cancel. As a step towards this, in this section, we prove some results about $R(W)$.

@@

That is, each $r_i$ is the determinant of the minor determined by the propagators (rows) that contribute different values (columns) to the $i^{th}$ element of the Grassmann necklace than they did to the $i-1^{th}$.

As we discuss the algorithm it will also be useful to have the notation $\Delta_{I_i}$ for the determinant of the $k\times k$ minor of $C(W)$ with columned indexed by $I_i$.

@@

@@

The irriducible quadratic factor corresponds to the fact that $q$ and $s$ share an edge {\em and} contribute both endpoints of that edge to $I_7$; see Proposition~\ref{prop alg gives rad} below.

@@

If neither $m$ nor $m+1$ appear in $I_i$ then by Corollary~\ref{no coloops} $V(p) \cap I_i = \{i\}$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$, which is a contradiction to the fact that $i$ contributes to $f$.  So we must have at least one of $m$ and $m+1$ in $I_i$.  However, all propagators in $S$ are mapped by the function $I_{i}(\cdot)$ to vertices strictly before $m$, so the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the $|S|\times |S|$ matrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. No other propagators can be supported on these vertices since all other propagators are outside of $p$, and $p$ is the first propagator supported at $i$; this explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$, and both factors are nontrivial since at least one of $m$ and $m+1$ appear in $I_i$.  Since $p\in \cP_{in}(p)$ we have $f|\det A$ and if we remove the propagator outside of $p$ that contributes $m$ or $m+1$, we get a diagram with fewer propagators for which $\det A | \Delta_{I_i}$.  This contradicts the minimality of our choices.

@@

Similarly to Case 1, let $S = \cP_{in}(q) \cup \{p\}$ in the $<_i$ order. Then all propagators in $S$ are mapped by $I_i(\cdot)$ to vertices strictly before $j$ and no other propagators are supported on vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $>_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}$.  This yields a contradiction as in Case 1: either $W$ contains unnecessary propagators which contradicts our minimality assumption, or $\det A$ is quadratic which contradicts the assumption that $p$ and $i$ contribute to $f$, an irreducible factor of degree at least 3.

@@

Now suppose two variables from the row $p$ appear in a quadratic factor $f$.  By part (3), there is another propagator $q$ and an edge $(a,b)$ such that $f$ is the $2\times 2$ minor coming from $p, q$ and $a, b$, with $I_i(p)=a$, $I_i(q)=b$.  There are two situations which can occur, both illustrated in Figure~\ref{fig quadratic}; we show that in both cases it follows that $q \in S_i$ as well.

@@

Now consider the case where $a<_i b$, and suppose for contradiction that $q \not\in S_i$, i.e. that $I_{i-1}(q) = b$. Since $I_{i-1}(p) \neq a$, there must be some other propagator $s$ with $I_{i-1}(s) = a$ (else $I_{i-1}$ assigns $q$ to $a$). This propagator cannot lie on edge $(a,b)$ since by Lemma~\ref{vertex cyclic int lem} we must have $I_i(s) = a$ or $b$, contradicting the fact that $I_i(p) = a$ and $I_i(q) = b$; thus $s$ has an end on $(a-1,a)$ and is inside $p$ from the point of view of $i-1$.

Say $s$ is supported on $(j, j+1, a-1, a)$ and $p$ is supported on $(k, k+1, a, b)$ with $i-1 \leq_{i-1} k+1 \leq_{i-1} j+1$. But by Lemma~\ref{lem no fourth vertex}, if $I_{i-1}(s) = a$ then $a$ cannot be maximal in the support of $s$ with respect to $<_{i-1}$; thus we must have $i-1 = j+1$, and we are in the situation in figure~\ref{fig part 4}.

@@

Since $p$ changed its association from $I_{i-1}$ to $I_i$, we have $I_{i-1}(p) = i-1$ by Lemma~\ref{vertex cyclic int lem}. From figure~\ref{fig part 4} it follows that $I_{i-1}$ assigns $p$ to $i-1$ and then proceeds identically to $I_i$ for all vertices inside $p$, implying that $I_{i-1}(s) = I_i(s)$. Since $I_{i-1}(s) = a$ and $I_i(s) \neq a$, this is a contradiction.

@@

\begin{proof}
The equivalence up to scalar multiple of $R$ and the radical of $\prod_{i=1}^n \Delta_{I_i}$ is due to Proposition~\ref{prop alg gives rad}.  It remains to prove that $R(W)$ is the $R$ of Algorithm~\ref{alg WLD to denom via GN}.

To this end, first note that $R(W)$ and $R$ both have total degree $4|\mathcal{P}|$; the degree of $R(W)$ is immediate from the definition while that of $R$ follows from Lemma~\ref{vertex cyclic int lem}.  By Proposition~\ref{prop alg gives rad} every factor of $R$ is either a single variable or a quadratic factor coming from two propagators supported on a common edge.  The factors of each $R_e$ making up $R(W)$ in the notation of Definition~\ref{def R(W)} are all of this form and hence every factor of $R$ divides $R(W)$.
Finally, since $R$ is squarefree, this implies that $R(W)$ is a scalar multiple of $R$.

@@

While the authors cannot yet offer a proof that the appropriate sum of the Wilson loop diagrams are, in fact, finite, as conjectured [cite], we do hope that this relationship to the Grassmann Necklace elements will provide a different avenue for answering that and related questions.

>>>>

Our next goal is to show that the dimension of the positroid cell defined by a Wilson loop diagram $(\cP, [n])$ has dimension $3|\cP|$ (Theorem \ref{thm dim}).  The positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite{non-orientable}, Agarwala and Marcott show that the dimension of the Deodhar component is 
\[|\cP| + \text{ the dimension of associated positroid cell}.\]
\note{What if instead of displaying this we wrote ``is the sum of $\|\cP\|$ and the dimension fo the associated positroid cell.''}
In other words, the Deodhar component associated to a Wilson loop diagram is $4|\cP|$, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron.

@@

\begin{proof}
  By Algorithm~\ref{alg:put GN on WLD} the Grassmann necklace of $W$ is obtained from the Grassmann necklace of $V$ by duplicating the $i$th element of the Grassmann necklace of $V$ (shifting indices by $1$ for all indices greater than $i$), and incrementing all indices greater than $i$ by $1$ in each Grassmann necklace element\note{This is saying to increment twice in a way which I think is confusing, but presumably the parenthetical remark was put there to try to clarify something so I don't want to just remove it}.  Formally, if $I_1^{V}, \ldots, I_{n-1}^{V}$ and $I_1^{W}, \ldots, I_n^{W}$ are the Grassmann necklaces of $V$ and $W$ respectively then
  \[
  I_j^{W} =
  \begin{cases}
    \{\ell \in I_j^{V} : \ell < i\} \cup \{\ell+1 \in I_j^{V} : \ell \geq i\} & \text{if $j\leq i$} \\
    \{\ell \in I_{j-1}^{V} : \ell < i\} \cup \{\ell+1 \in I_{j-1}^{V} : \ell \geq i\} & \text{if $j > i$.}
  \end{cases}
  \]

By Lemma~\ref{no coloops} we know that $i \not\in I_1^W$, and so $i$ must label a horizontal edge on the boundary of the Le diagram of $W$, i.e. it must be a column label. The shapes of the Le diagram of $V$ and $W$ are the same except for the insertion of this column since $I_1^*$\note{I don't find the star very helpful here; different area's conventions maybe} is the same for $V$ and $W$ except for the incrementation of the indices $\geq i$ in the transition from the necklace for $V$ to the necklace for $W$. 
\end{proof}

@@

We first consider $I_1^{W}$, i.e. the set obtained by applying the Grassmann necklace algorithm to $W$ with starting vertex 1\note{I don't find this i.e. helpful at this point, is it an old thing that can go or do others find it useful?}. Note that $n-3 \not\in I_1^{V}$: for Case 2 this is clear from the diagram (since Algorithm~\ref{alg:put GN on WLD} would assign $q$ no later than vertex $n-4$), while for Case 1 it follows from Lemma \ref{lem no fourth vertex}. Therefore when we start at vertex 1 and apply the Grassmann necklace algorithm to $W$, $p$ is the only remaining unassigned propagator when we reach vertex $n-3$, and so we have $ I_1^{W}(p) = n-3$ and $I_1^W = I_1^V\cup \{n-3\}$.

@@

Finally, if $n-2,n-1 \in I_{n-2}^{V}$ then $I_{n-2}^W(q) = n-2$ and $I_{n-2}^W(p) = n-1$ as above, but this is different to what occured in $I_{n-2}^{V}$ so we cannot use the same argument as above. When the algorithm $I_{n-2}^W$ reaches vertex $n$, only the propagators $p$ and $q$ have contributed; thus we proceed as in the construction of $I_n^{V}$ but without propagator $q$. By Lemma~\ref{vertex cyclic int lem} we know that $I_n^V(q) = n-5$, and from the diagram we see that the only way this could occur is if all other propagators in $V$ had already been assigned when we reached vertex $n-5$; thus $q$ was the final propagator to be assigned in $I_n^{V}$. Therefore $I_{n-2}^{W} = (I_{n}^{V} \setminus \{n-5\}) \cup \{n-1, n-2\}$ in this case. This completes all cases for $I_{n-2}^{W}$.

@@

As illustrated in Figure~\ref{fig Le}, the pieces of the Le diagrams of $V$ and $W$ will be called $\mathcal{A}$ and $\mathcal{B}$ in what follows. \hlfix{Explicitly, $\mathcal{B}$ is the set of columns with labels $v < n-3$ and $\mathcal{A}$ is the set of columns labeled by $n$, $n-1$ and $n-2$ in $V$. In $W$, $\mathcal{A}$ consists of these columns without the last row.}{is this accurate?}\note{I think it is accurate} Over the course of the next few lemmas we will prove that the plusses in the $\mathcal{B}$ parts of each diagram are identical, and that the relationship between the plusses in the $\mathcal{A}$ regions can be described explicitly.

@@

Finally, we also note that when we speak of a plus in the Le diagram of $V$ being the same as in $W$ or vice versa, we mean that the position of the plus in $\mathcal{A}$ or $\mathcal{B}$ is the same.

@@

In this section, we investigate the poles of the integrand for $I(W)$, that is $R(W)$ as defined in Definition~\ref{def R(W)}. It is conjectured that for a fixed number of propagators and a given set of external twistors $\cZ$, [cite]\note{A cite still waiting} \bas \sum_{W | |\cP| = k} I(W)(\cZ) <\infty \;.\eas In otherwords, all poles of the intergrals must cancel. As a step towards this, in this section, we prove some results about $R(W)$.

@@

That is, each $r_i$ is the determinant of the minor determined by the propagators (rows) that contribute different values (columns) to the $i^{th}$ element of the Grassmann necklace than they did to the $i-1^{st}$.

As we discuss the algorithm it will also be useful to have the notation $\Delta_{I_i}$ for the determinant of the $k\times k$ minor of $C(W)$ with columns indexed by $I_i$.

@@

\note{I would like to discuss the angles of the arrows in fig 14}

@@

The irreducible quadratic factor corresponds to the fact that $q$ and $s$ share an edge {\em and} contribute both endpoints of that edge to $I_7$; see Proposition~\ref{prop alg gives rad} below.

@@

If neither $m$ nor $m+1$ appear in $I_i$ then by Corollary~\ref{no coloops} $V(p) \cap I_i = \{i\}$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$, which is a contradiction to the fact that $i$ contributes to $f$.  So we must have at least one of $m$ and $m+1$ in $I_i$.  However, all propagators in $S$ are mapped by the function $I_{i}(\cdot)$ to vertices strictly before $m$, so the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the $|S|\times |S|$ matrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. No other propagators can be supported on these vertices since all other propagators are outside of $p$, and $p$ is the first propagator supported at $i$; this explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$, and both factors are nontrivial since at least one of $m$ and $m+1$ appear in $I_i$.  Since $p\in \cP_{in}(p)$ we have $f|\det A$ and if we remove the propagator outside of $p$ that contributes $m$ or $m+1$, we get a diagram with fewer propagators for which $\det A | \Delta_{I_i}$, hence $f|\Delta_{I_i}$.  This contradicts the minimality of our choices.

@@

Similarly to Case 1, let $S = \cP_{in}(q) \cup \{p\}$ in the $<_i$ order. Then all propagators in $S$ are mapped by $I_i(\cdot)$ to vertices strictly before $j$ and no other propagators are supported on vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $>_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}=\det A\det B$.  This yields a contradiction as in Case 1: $f|\det A$ and removing a propagator not in $S$ gives a smaller diagram with a factor of degree $\geq 3$ which contradicts our minimality assumption.\note{I rephrased this last sentence quite a bit because earlier things had gotten rephrased}

@@

Now suppose two variables from the row $p$ appear in a quadratic factor $f$ of $\Delta_{I_i}$.  By part (3), there is another propagator $q$ and an edge $(a,b)$ such that $f$ is the $2\times 2$ minor coming from $p, q$ and $a, b$, with $I_i(p)=a$, $I_i(q)=b$.  There are two situations which can occur, both illustrated in Figure~\ref{fig quadratic}; we show that in both cases it follows that $q \in S_i$ as well.

@@

Now consider the case where $a<_i b$, and suppose for contradiction that $q \not\in S_i$, i.e. that $I_{i-1}(q) = b$. Since $I_{i-1}(p) \neq a$, there must be some other propagator $s$ with $I_{i-1}(s) = a$ (else $I_{i-1}$ assigns $q$ to $a$). This propagator cannot lie on edge $(a,b)$ since, by Lemma~\ref{vertex cyclic int lem}, we must have $I_i(s) \in \{a, b\}$, contradicting the fact that $I_i(p) = a$ and $I_i(q) = b$; thus $s$ has an end on $(a-1,a)$ and is inside $p$ in the $<_k$ order where $p=(k,a)$.

Say $s = (j,a)$ with $i-1 \leq_{i-1} k+1 \leq_{i-1} j+1$. But by Lemma~\ref{lem no fourth vertex}, if $I_{i-1}(s) = a$ then $a$ cannot be maximal in the support of $s$ with respect to $<_{i-1}$; thus we must have $i-1 = j+1$, and we are in the situation in Figure~\ref{fig part 4}.

@@

Since $p$ changed its association from $I_{i-1}$ to $I_i$, we have $I_{i-1}(p) = i-1$ by Lemma~\ref{vertex cyclic int lem}. From Figure~\ref{fig part 4} it follows that $I_{i-1}$ assigns $p$ to $i-1$ and then proceeds identically to $I_i$ for all vertices inside $p$, implying that $I_{i-1}(s) = I_i(s)$. Since $I_{i-1}(s) = a$ and $I_i(s) \neq a$, this is a contradiction.

@@

\begin{proof}
The equivalence of the ideal generated by $R$ and the radical of the ideal generated by $\prod_{i=1}^n \Delta_{I_i}$ is due to Proposition~\ref{prop alg gives rad}.  It remains to prove that $R(W)$ is the $R$ of Algorithm~\ref{alg WLD to denom via GN}.

To this end, first note that $R(W)$ and $R$ both have total degree $4|\mathcal{P}|$; the degree of $R(W)$ is immediate from the definition while that of $R$ follows from Lemma~\ref{vertex cyclic int lem}.  By Proposition~\ref{prop alg gives rad} every factor of $R$ is either a quadratic factor coming from two propagators supported on a common edge or a single variable coming from a support vertex of a propagator not appearing in a quadratic factor.  The factors of each $R_e$ making up $R(W)$ in the notation of Definition~\ref{def R(W)} are all of this form and every such factor occurs.  Hence every factor of $R$ divides $R(W)$.
Finally, since $R$ is squarefree, this implies that $R(W)$ is a scalar multiple of $R$.

@@

While the authors cannot yet offer a proof that the appropriate sum of the Wilson loop diagrams are, in fact, finite, as conjectured [cite]\note{One last cite to fill out}, we do hope that this relationship to the Grassmann Necklace elements will provide a different avenue for answering that and related questions.
